---
title: 'Stat 480 - Homework #2'
author: "Your Name"
date: "1/30/2020"
output: html_document
---
---
title: 'Stat 480 - Homework #2'
author: "Your Name"
date: "1/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Economic Guide to Picking a College Major

1. Download the RMarkdown file with these homework instructions to use as a template for your work. Make sure to replace "Your Name" in the YAML with your name.

2. Using the command below, read in the data set into your R session. 

```{r}
recent_grads <- read.csv("https://raw.githubusercontent.com/Stat480-at-ISU/materials-2020/master/02_r-intro/data/recent_grads.csv")
```

3. Create a new variable, `share_women`,  in the dataset that is women as share of total; i.e. the number of women divided by the total number of men and women.

4. Create a subset of the data containing only the rows where the `Major_category` is `STEM`.

5. For your subset, compute the average  of `share_women` and its standard deviation. Also compute the mean median earnings (`Median`) and its standard deviation. Comment on the results. (You might have to deal with missing values appropriately).

5. Again using the subset, compute the correlation between women as a share of total (`share_women`) and the median earnings (`Median`) and interpret your results.

6. Use the original dataset and `ggplot2` to draw a scatterplot of women as share of total and the median earnings. Color points by the major category (`Major_category`). Comment on the result. 



**Due date**: the homework is due before class on Thursday. 

**For the submission**: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html file with it. 
