---
title: 'Stat 480 - Homework #3'
author: "Your Name"
date: "2/6/2020"
output: html_document
---
---
title: 'Stat 480 - Homework #3'
author: "Your Name"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spotify data

1. Download the RMarkdown file with these homework instructions to use as a template for your work. Make sure to replace "Your Name" in the YAML with your name.  

2. Using the command below, read in the spotify data set into your R session. 

```{r}
spotify <- read.csv("https://raw.githubusercontent.com/Stat480-at-ISU/Stat480-at-ISU.github.io/master/homework/data/spotify.csv")
```

3. Use one of our object inspecting functions and interpret the result in the data that you see.

4. Use the package `ggplot2` to draw a barchart of the genres. In addition, map the genre categories to the fill color of the barchart.

5. Use the package `ggplot2` to draw a histogram of one of the continuous variables in the dataset. Use fill color to show the genre categories and adjust the binwidth if necessary. Use `facet_wrap()` to create a histogram for each of the genre categories.

6. Use the package `ggplot2` to draw a boxplot to compare one of the continuous variables with the genre categories. Again, use fill color to show the genre categories.

7. For each of the three figures above, write a two-three sentence summary, describing the
    a. **structure of the plot**: what type of plot is it? Which variables are mapped to x, to y, and to the (fill) color?
    b. **main message of the plot**: what is your main finding, i.e. what do you want viewers to learn from the plot?
    c. **additional message**: point out anomalies or outliers, if there are any. 


Due date: the homework is due before class on Thursday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html file with it. 
