---
title: "Solution to Sample Midterm Stat 480"
author: "Heike Hofmann"
date: "April 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

Work on questions in R, make sure to keep a copy of your R code - you will be asked to submit this script at the end.

Data on all flights in and out of Des Moines (DSM) for October 2008 are available at <a href="http://www.hofroe.net/data/dsm-flights.csv">http://www.hofroe.net/data/dsm-flights.csv</a>.<br>
See <a href="http://stat-computing.org/dataexpo/2009/the-data.html">http://stat-computing.org/dataexpo/2009/the-data.html</a> for a description of the variables.

1. Load the flights data into R. <br>
		Determine, which flight was delayed the most on arrival. Report its row number, where it started, and by how much the flight was delayed on departure.  
		
		
```{r}
flights <- read.csv("http://www.hofroe.net/data/dsm-flights.csv")
which.max(flights$ArrDelay)
flights[which.max(flights$ArrDelay), c("Origin", "DepDelay")]
```

2. Bring the variable 'Day' into the correct order, starting with 'Monday'.

```{r}
summary(flights$Day)
days <- levels(flights$Day)
flights$Day <- factor(flights$Day, levels=days[c(2,6,7,5,1,3,4)])
summary(flights$Day)
```

3. Create a new variable called 'Weekend' which has value TRUE for Saturdays and Sundays and FALSE otherwise. 

```{r}
# create new variable Weekend
flights$Weekend <- flights$Day %in% c("Saturday", "Sunday")

summary(flights$Weekend)
```

4. Determine how many flights arrived in Des Moines on average each day of the week. 

```{r}
# idea 1:
table(subset(flights, Dest=="DSM")$Day) # overall number of flights by day of week
# problem: how many Mondays, Tuesdays, are there in October 2008?

require(lubridate)
octs <- data.frame( date = ymd(paste("2008/10/",1:31, sep="")))
octs$day = wday(octs$date, label=TRUE)
table(octs$day)

table(subset(flights, Dest=="DSM")$Day)/c(4,4,5,5,5,4,4)

# idea 2: 
require(dplyr)
require(lubridate)
flights %>% filter(Dest == "DSM") %>% group_by(DayofMonth) %>% summarise(
  day = Day[1],
  n = n()
) %>% group_by(day) %>% summarize(avg = mean(n), n = n())
```

6. How many flights were scheduled to go to Denver (DEN)? What percentage of flights goes to Denver? 

```{r}
nrow(subset(flights, Dest=='DEN'))
nrow(subset(flights, Dest=='DEN'))/nrow(subset(flights, Dest != 'DSM'))*100
```

7. Where do most flights arriving in Des Moines come from? (use IATA code) </p>

```{r}
flights %>% 
  filter(Origin != "DSM") %>% 
  group_by(Origin) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))

sort(table(flights$Origin), decreasing=T)[2]
```

8. Plot boxplots of arrival delays by originating airports. Order boxplots according to increasing median arrival delay. </p>

```{r}
library(ggplot2)
flights %>% filter(Dest =="DSM") %>%
ggplot(	aes(x = reorder(factor(Origin), ArrDelay, na.rm=T), y = ArrDelay)) + geom_boxplot()
```

9. Using dplyr, determine for flights leaving DSM for each hour of the day 

  - the number of flights scheduled to leave, 
  - the percentage of flights departing late (FAA considers flights late, if they are 15 minutes or more behind the scheduled time), 
  - the average departure delay, and 
  - the top 3 destinations.

Draw a scatterplot of average departure delay by scheduled hour of departure. Color points by top destination, adjust point size to reflect the number of flights for each hour. 

```{r}
dep.summary <- flights %>% 
  filter(Origin == 'DSM') %>%
  mutate(hour = CRSDepTime%/%100) %>% 
  group_by(hour, Dest) %>%
  mutate(test = n()) %>% 
  group_by(hour) %>% 
  mutate(rank = dense_rank(desc(test))) %>% 
  summarise(
    count = n(),
    pct.delayed = sum(DepDelay>15, na.rm=TRUE)/n()*100,
    avg.delay = mean(DepDelay, na.rm=T),
    top.Dest.1 = Dest[which.max(test)],
    top.Dest.2 = first(Dest[which(rank == 2)]),
    top.Dest.3=first(Dest[which(rank == 3)]))


dep.summary %>% 
  ggplot(aes(x = hour, avg.delay, colour = top.Dest.1, size = count)) + geom_point()
```

